\textbf{Title}

### 2015-06-01 paper/draft_3.tex
some text revision

In recent years, the use of various shrinkage and thresholding procedures has become ever more popular in a variety of applications where sparsity is a desirable notion. Here the term ``sparsity'' is loosely defined in the sense that it also includes the case where there are many near-zero (as opposed to exactly zero) terms, as shrinkage procedures often produce. The most prominent usage of such procedures appear in multiresolution denoising analyses such as wavelets, where sparsity in a transformed domain often corresponds to smoothness in the data space [references]. Other examples include estimation of covariance and precision matrices (though thresholding is used more often here since zero terms are usually desired) [references], accounting for different measurement precisions in FDR control for multiple testing [references], or more generally optimizing penalized likelihood problems such as LASSO [references]. In this paper we will look at a flexible and adaptive shrinkage method named ASH, and illustrate its value in two key areas of univariate wavelet denoising.\bigskip\\
Before delving deeper into the two denoising tasks, we first present a brief overview of wavelet estimators. These estimators have become extremely popular in nonparametric regression problems following the seminal paper by Donoho \& Johnstone (1994). Not only are wavelet methods locally adaptive in the sense that they achieve the optimal minimax rate over a wide variety of functions, but they are also computationally faster than many other adaptive methods such as variable-bandwidth kernel methods [antoniadis 2001]. While classical thresholding estimators have been shown to possess optimal asymptotic properties, an extensive simulation study by Antoniadis et al. (2001) has demonstrated that various Bayesian shrinkage methods are usually more data adaptive than thresholding methods in finite samples, outperforming the latter in terms of mean squared error (MSE). However, many existing Bayesian wavelet methods are often computationally intensive, and may not be practical for large scale applications. Furthermore, it is not easy to apply the same thresholding or Bayesian shrinkage rules to other nonparametric regression problems, two of which we subsequently describe. On the other hand, our method is computationally faster than previous Bayesian methods while still possessing good finite sample performance, and uses ASH as the common shrinkage procedure in both the problems.\bigskip\\
The first problem relates to the frequently encountered Gaussian case. Although the theory for mean estimation with homoskedastic Gaussian errors has been thoroughly developed in the wavelet literature [references], the existing methods for variance estimation with heteroskedastic errors have been far and few between. Fan \& Yao (1998) estimated the variance by smoothing the squared residuals using local polynomial smoothing, while Brown \& Levine (2007) employed difference-based kernel estimators. Making use of the local adaptivity of wavelet methods, Cai \& Wang (2008) improved upon previous approaches using a wavelet thresholding approach on first order differences. While Cai \& Wang (2008) proved that their technique was asymptotically optimal, we will demonstrate how we can further improve the finite sample performance of previous variance estimation procedures by incorporating ASH into a novel wavelet denoising framework. [mention the lack of available software?]\bigskip\\
The second task is to denoise a Poisson distributed signal. This often occurs in the experimental sciences such as gamma-ray burst signals in astronomy [references], and more recently in genetics where high throughput sequencing data is of interest. This problem is particularly interesting in the sense that there are many different ways to perform signal recovery. Variance stabilizing techniques together with normal approximations have been proposed by Donoho (1993) and Fryzlewicz \& Nason (2001), by noting that the variance of a Poisson count is equal to its mean. Kolaczyk (1997, 1999a) derived thresholds achieving optimal asymptotic properties in the context of wavelet transformations, similar to the thresholds in the i.i.d. Gaussian case.  However, variance stabilizing methods are computationally inefficient due to the presence of external cycle-spinning, and threshold-based methods may not result in optimal finite sample performance. Furthermore, both these methods are quite sensitive to the choice of the primary resolution level due to the Poisson nature of the data. Multiscale analysis using recursive dyadic partitions within a Bayesian framework was later developed by Kolaczyk (1999b) to make use of a particular form of likelihood factorization, but a relatively inflexible conjugate prior was chosen. We will improve upon the prior (and likelihood) specification by using ASH as the main shrinkage procedure in a Bayesian framework similar to that of Kolaczyk (1999b). \bigskip\\
Both these problems are generally harder to deal with than the classical problem of i.i.d. Gaussian errors, and usually requires some work in extending the ideas from methods aimed at the latter task. Before proceeding to describe the two problems in detail, we will first describe briefly the portion of ASH that performs shrinkage. \bigskip\\

@@

Note that $\bm{\pi}$ and $\bm{s}^2$ is shared across all the $\Gb_i$'s, allowing one to borrow information across all the observations. The full likelihood for $\bm{\pi}$ given $\bm{\hat{\Gb}}$ and $\bm{\hat{\s}}$ is then a product of likelihoods for $\bm{\pi}$ given each $\hat{\Gb}_i$ and $\hat{\s}_i$. Since the likelihoods are Gaussian and the priors are mixtures of Gaussians, the posterior distribution of $\Gb_i$ given $\hat{\Gb}_i$ and $\bm{\hat{\pi}}$ is analytically tractable, where the mixing proportions $\bm{\hat{\pi}}$ are maximum likelihood estimates. The posterior mean of $\Gb_i$ is taken to be the point estimate of $\Gb_i$, resulting in a shrinkage procedure.\bigskip\\
The flexibility and computational efficiency of ASH is immediately obvious, which motivates us to extend this method to the more complex wavelet setting described here. As with many other Bayesian procedures, ASH borrows information across all available data, and is hence extremely data adaptive. At the same time, the key advantage of ASH over other Empirical Bayes procedures such as EbayesThresh [Johnstone \& Silverman (2005)] and the commonly used ``spike-and-slab'' prior (is there a name for this?) is that it allows for variable variances as input, a feature that the latter methods lack. This is also the primary reason we can use ASH as the only shrinkage procedure in both the problems described above. Furthermore, Gaussian mixture priors can effectively approximate any unimodal distribution, allowing for a more flexible prior than many other procedures. At the same time, we can speed up the entire signal denoising problem via likelihood approximations (discussed in the next section), allowing our method to be computationally efficient without compromising too much accuracy. It should be noted that typical wavelet methods require the specification of a primary resolution level as an extra parameter, which could substantially influence the accuracy of the method. However, we are able to apply ASH to every resolution in the wavelet transformation, sparing the need for such a parameter. For an in-depth understanding of the original motivations and applications of ASH, see Stephens (??). With the main shrinkage method accounted for, we will now proceed to explore the two aforementioned problems in detail. Note that one only needs to supply a vector of estimates $\hat{\Gb}$ and their standard errors se$(\hat{\Gb})$ to ASH to obtain posterior estimates; hence, we will focus on obtaining these estimates and standard errors when describing our methods.

@@

with $\sum_l \pi_l^{(j)}=1$, for $j=1,...,J$. $\pi_i^{(j)}$ and $(s_i^{(j)})^2$ are hyperparameters that are shared between coefficients in the same resolution level, and $m$ is the number of mixture components. In ASH, $m$ is usually chosen so that it is large enough to reasonably approximate any unimodal distribution, while maintaining a reasonable computational speed. Note that, for notational convenience, we have switched to a double index following standard wavelet convention for indices: here $j=1,...,J$ is the resolution level, and $k=0,...,2^j-1$ is the location within each resolution level $j$. By applying the inverse wavelet transform on the posterior mean $\hat{\bm{\Ga}}$ that ASH produces, one can obtain the posterior mean of $\bm{\mu}$, which serves as an estimate that minimizes the MSE. One could also construct credible bands using the posterior variances.\bigskip\\
While many existing methods do a good job of signal denoising in the presence of i.i.d. (Gaussian) errors, heteroskedastic (but still independent) errors present a different challenge. One key obstacle when dealing with heteroskedastic errors is that the likelihood for $\bm{\Ga}$ does not necessarily factorize. If the true variances were known, one could obviously write out the full likelihood and proceed to compute the posterior mean/median using some specified prior. However, this would be computationally cumbersome, and may not be feasible when extended to multiple signals or higher dimensions. A suitable prior might also be difficult to find. As such, one key aspect of our approach is to treat the wavelet coefficients as if they were independent, so that the true likelihood is approximated by a composite likelihood (see Silverman (1999) for an example where this is done).\bigskip\\

@@

where $\Sigma=\textrm{diag}(\s_1,...,\s_T)$ is a diagonal matrix. For the rest of this subsection, we will use the non-decimated wavelet transform (NDWT) instead of the standard DWT to achieve translation invariance. This property of translation invariance reduces the presence of artifacts, which often occur near discontinuities in the underlying signal (see eg. Coifman \& Donoho (1995)). Now let $W$ be the matrix associated with the NDWT for a given wavelet basis, so that

@@

where $P(d_{jk}|\Ga_{jk})=\phi(d_{jk};\Ga_{jk},\tilde{\Sigma}_{(jk,jk)})$. Note that there are $n$ coefficients at each resolution level instead of $2^j$ for resolution $j$ because we are using the NDWT instead of the standard wavelet transform. Here $\phi$ denotes the Gaussian density function. Since $\Sigma$ is diagonal, it is easy to see that $\tilde{\Sigma}_{(jk,jk)}=\sum_{i=1}^T \Sigma_{ii}W_{jk,i}^2$. In our method, we use ASH to assign independent priors to $\Ga_{jk}$ as with \eqref{eq:ashprior}. %ASH also allows us to estimate the $\pi_i$'s via a maximum likelihood approach and fix the $\s_i$'s by using a grid of values. This particular choice of prior specification is extremely flexible, being able to approximate any symmetrical zero-centered distribution, as well as allowing for a relatively easy estimation of the hyperparameters through empirical Bayes. See ASH for more details regarding this choice of prior (?)

@@

While we assumed that the true variances are known for mean estimation, the problem of variance estimation itself is a non-trivial one. Here we make the reasonable assumption that the variance function is also spatially structured, in addition to the mean function. The approach we took is similar in spirit to that of Cai \& Wang (2008), since we make use of wavelet decomposition as well. However, while they use first order differences, we look at the squared residuals. As such, we need a reasonably accurate estimator of the mean function in order to form sensible residuals. For clarity of presentation, we will first assume that the true mean function is known. Define

@@

Now that we can now estimate $\bm{\mu}$ given $\bm{\s}^2$ and vice versa, a natural procedure for jointly estimating $\bm{\mu}$ and $\bm{\s}^2$ is an iterative one. That is, we estimate $\bm{\mu}$ given some initial estimate of $\bm{\s}^2$, then estimate $\bm{\s}^2$ given the previous estimate of $\bm{\mu}$, and repeat this procedure for as long as is necessary. Although this iterative procedure (using squared residuals) could go on for a long time, and convergence is difficult to prove, we found via simulations that two cycles is usually sufficient to produce relatively accurate estimates. A natural initial estimate would be the square of first order differences between adjacent points, as discussed in previous variance estimation papers (eg Cai \& Wang (2008)). This estimator has the property that it is approximately unbiased for the variance at the two corresponding points, provided that the mean and the variance function is smooth enough. To be more specific, the initial variance estimates are defined by

@@

where $Y_0\equiv Y_n$ and $Y_{n+1}\equiv Y_1$, since wavelet methods typically assume that the function is periodic. In summary, our approach for joint mean and variance estimation can be described in three steps:

@@

\item Using squared first order differences \eqref{eq:initial var est} as our estimates of $\bm{\s}^2$, we estimate $\bm{\mu}$ as if $\bm{\s}^2$ was known (see the above procedure for the case when $\bm{\s}^2$ is known)
\item Given the estimate of $\bm{\mu}$, take squared residuals and treat those as ``observations'' for $\bm{\s}^2$. Next, we project them into wavelet space and apply some form of shrinkage to the wavelet coefficients, before projecting them back into data space to obtain an estimate of $\bm{\s}^2$.

@@

Next we turn to another popular nonparametric regression problem and demonstrate how we could make use of ASH. Specifically, we assume an underlying signal $\Gl_i$, $i=1,...,n$ with $n=2^J$ for some positive integer $J$ as with the Gaussian case. Further assume that each data point $Y_i$, $i=1,...,n$ is realized from a Poisson distribution with mean $\Gl_i$ ie. $Y_i=\textrm{Pois}(\Gl_i)$. Our goal is to recover the true intensity $\bm{\Gl}$ as accurately as possible, given $Y_k$. To make our approach easier to understand and relate it to wavelet-based methods, we first summarize the data in a recursive manner (see Kolaczyk (1999) and Nowak (1998):

@@

for resolution $j=0,...,J-1$ and location $k=0,...,2^j-1$, so that we are looking at the sum of adjacent pairs of ``points'' as we move to coarser resolutions, where a point is already itself a sum of the true data points. Another way to define the $Y$'s is

@@

where the factorization is due to the recursive definition above. Note that $Y_{00}|\Gl_{00}\sim \textrm{Pois}(\Gl_{00})$. Now for any give $j,k$, $Y_{jk}$ is a sum of two independent Poisson random variables, and is itself a Poisson random variable. Hence

@@

Whereas Kolaczyk (1999) sets a conjugate prior (which is a mixture of a point mass at 0.5 and a uniform distribution, similar to the ``spike and slab'' prior for the Gaussian case) for the parameters $\frac{\Gl_{j+1,2k}}{\Gl_{jk}}$ and uses a Binomial likelihood, we felt that the two-component mixture prior is too inflexible. As such, we work with the parameters $\Ga_{sl}$ as defined in \eqref{eq:poisson wc} and approximate the logit-Binomial (is there such a thing?) likelihoods by Gaussian likelihoods. Furthermore, these parameters can be more easily connected to the true wavelet coefficients since they consist of (log) differences rather than ratios.\bigskip\\

@@

where $S$ is the number of successes in the binomial setup, and $F=N-S$ is the number of failures (in our case the even locations would be successes and odd locations failures). The MLE has one notable drawback however, in that it cannot deal well with extreme cases. This would in turn affect the likelihood approximation in a non-negligible way. To improve the accuracy of the approximation at the endpoints of the binomial distribution (ie. when we have $\hat{p}=0$ or $N$, so that $\hat{\Ga}=-\infty$ or $\infty$ respectively), we instead aim to find a suitable estimator $\hat{\Ga}$ for $\Ga$. Next, we approximate the distribution of $\hat{\Ga}$ by a Gaussian distribution, so that the likelihood for $\Ga$ is automatically Gaussian, given by $\phi(\Ga;\hat{\Ga},\mathbb{V}(\hat{\Ga}))$. To that end, our estimator will still be the MLE, but with Tukey's modification (Gart \& Zweifel (1967)) at the endpoints where $S=\{0,N\}$. Specifically, our final estimators for $\Ga$ and its standard error are given by

@@

Finally, to reconstruct the signal in the original space using the posterior distribution of the $\Ga$'s (which are in the logit space), we make use of the Delta method. For full details on the reconstruction process as well as the Delta method, see Appendix \ref{app:reconstruction}.

@@

In all the scenarios presented in Tables \ref{table:hetero_v2_1} - \ref{table:hetero_bump_3}, it is obvious that the mean estimates deviate substantially from the true mean if the heteroskedastic nature of the variance functions is not taken into account. We also note that our method can nearly achieve the ``optimal'' MISE (ie the MISE for the case when our method is given the true variance) for variance function $V_2$, but not for the other. This is likely due to the fact that function $V_2$ is relatively smooth, and hence easier to estimate than the extremely jumpy and peaky Bumps function. Furthermore, the MISEs of our method are often smaller than the ones produced by TI-thresholding even when TI-thresholding is supplied the true variance function, where the latter procedure has been shown to be asymptotically optimal (see Gao (1997)). Hence, the finite sample performance of our method is guaranteed to a certain extent through these simulation studies. At the same time, using the variance function estimated by our method instead of running MAD also improves the performance of TI-thresholding, demonstrating the ability of our method to estimate the variance function relatively accurately. However, our method is still superior to TI-thresholding no matter the type of variance estimate that TI-thresholding uses in these simulation studies.\bigskip\\

@@

In the first scenario, the number of data points is not a power of two, nor are the point equally spaced. To deal with these complications, we adapted and modified the standard symmetric extension procedure commonly adopted in wavelet settings. We first mirrored the data about the right edge and extract the first $2^{\lfloor\log_2(2n)\rfloor}$ sample points. This ensures that the number of data points in the new ``dataset'' is a power of two, and the mean curve would be continuous at the right edge. To further ensure that the input to the Gaussian denoising method is periodic, we then reflected the new dataset about the right edge and used this as the final input. In order to obtain the original mean and variance functions, we extracted the first $n$ points from the outputs (mean and variance) of our denoising technique. Since the data points are not evenly spaced, we took the simplest approach and applied our method treating the observations as if they were evenly spaced. This approach is not only intuitively appealing, but can also be considered a formal treatment of unequally spaced data in traditional wavelet settings (see Sardy et al. (1999)). Evaluation of MSE at the 201 equally spaced points is then based on simple linear interpolation between the estimated points. Tables \ref{table:mfvb_comp} displays the MSEs over 100 independent runs for each scenario.

@@

begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & (0.01,3) & (1/8,8) & (1/128,128) \\
  \hline
SMASH & 1.00 & 1.00 & 1.00 \\
  BMSM & 1.55 & 1.58 & 1.21 \\
  BMMIM & 1.32 & 1.86 & 2.06 \\
  Haar-Fisz & 1.20 & 1.39 & 1.31 \\
  Anscombe, universal & 1.55 & 1.88 & 1.39 \\
  Haar thresholds & 1.35 & 1.38 & 1.23 \\
   \hline
\end{tabular}
\caption{Relative MSE of various methods for the Bumps function for Poisson data}
\label{table:pois_bumps}
\end{table}

@@

In this paper we have briefly introduced the adaptive shrinkage method ASH; while it was originally developed in the setting of FDR control for multiple comparisons, we have illustrated its usage as part of two wavelet denoising techniques. Both the applications discussed in this paper relaxes the standard assumption of i.i.d. Gaussian noise, and thus present us with challenging tasks. Through these applications we are able to demonstrate the flexibility and accuracy of the shrinkage method, allowing it to be potentially used in a variety of other applications where shrinkage is desirable.\bigskip\\

>>>>

In recent years, various shrinkage and thresholding procedures have become ever more popular in many applications where sparsity is desired. Here ``sparsity'' is loosely defined in that it also includes the case where many terms are near-zero, as opposed to exactly zero. The most prominent usage of such procedures appear in multiresolution denoising analyses such as wavelets, where sparsity in a transformed domain often corresponds to smoothness in the data space [references]. Other examples include estimation of covariance and precision matrices (though thresholding is used more often here since zero terms are actually desired) [references], accounting for different measurement precisions in FDR control for multiple testing [references], or more generally optimizing penalized likelihood problems such as LASSO [references]. In this paper we will look at a flexible and adaptive shrinkage method named ASH, and illustrate its value in two key areas of univariate wavelet denoising.\bigskip\\
Before delving deeper into the two denoising tasks, we first present a brief overview of wavelet estimators. These estimators have become extremely popular in nonparametric regression problems following the seminal paper by Donoho \& Johnstone (1994). Wavelet methods are locally adaptive in the sense that they achieve the optimal minimax rate over a wide variety of functions, yet are also computationally faster than many other adaptive methods such as variable-bandwidth kernel methods [antoniadis 2001]. While classical thresholding estimators have been shown to possess optimal asymptotic properties, an extensive simulation study by Antoniadis et al. (2001) has demonstrated that various Bayesian shrinkage methods are usually more data adaptive than thresholding methods in finite samples, outperforming the latter in terms of mean squared error (MSE). However, many Bayesian methods require specific distributional assumptions, and are not easily extensible to more complex problems, two of which we subsequently describe. On the other hand, our method simplies the task by making approximations, while still possessing good finite sample performance and computational efficiency. These approximations also allows us to use ASH as the common shrinkage procedure in both the problems.\bigskip\\
The first problem relates to mean and variance estimation with Gaussian noise. Although the framework for homoskedastic Gaussian errors has been thoroughly developed in the wavelet literature [references], existing methods dealing with heteroskedastic errors have been far and few between. Fan \& Yao (1998) estimated the variance by smoothing the squared residuals using local polynomial smoothing, while Brown \& Levine (2007) employed difference-based kernel estimators. Making use of the local adaptivity of wavelet methods, Cai \& Wang (2008) improved upon previous variance estimation methods using a wavelet thresholding approach on first order differences. Here we present an approach that estimates both the mean and variance accurately by incorporating ASH into a novel wavelet denoising framework. [mention the lack of available software?]\bigskip\\
The second task is to denoise a Poisson distributed signal. This often occurs in the experimental sciences such as gamma-ray burst signals in astronomy [references], and more recently in genetics where high throughput sequencing data is of interest. This problem is interesting in that there are many different ways to perform signal recovery. Variance stabilizing techniques together with normal approximations have been proposed by Donoho (1993) and Fryzlewicz \& Nason (2001), by exploiting the mean-variance relationship of a Poisson distribution. Kolaczyk (1997, 1999a) derived thresholds achieving optimal asymptotic properties in the context of wavelet transformations, similar to the thresholds in the i.i.d. Gaussian case.  However, variance stabilizing methods are computationally inefficient due to the presence of external cycle-spinning, and threshold-based methods may not result in satisfactory finite sample performance. Furthermore, both these methods are quite sensitive to the choice of the primary resolution level as simulations have shown. Multiscale analysis using recursive dyadic partitions within a Bayesian framework was later developed by Kolaczyk (1999b) to make use of a particular form of likelihood factorization, but a relatively inflexible conjugate prior was chosen. We will improve upon the prior (and likelihood) specification by using ASH as the main shrinkage procedure in a Bayesian framework similar to that of Kolaczyk (1999b). \bigskip\\
Both the aforementioned problems are harder to deal with than the classical problem of i.i.d. Gaussian errors, and usually requires one to extend the ideas from the latter task. Before proceeding to describe the two problems in detail, we will first describe briefly the portion of ASH that performs shrinkage. \bigskip\\

@@

Note that $\bm{\pi}$ and $\bm{s}^2$ are shared across all the $\Gb_i$'s, allowing one to borrow information across all observations. The full likelihood for $\bm{\pi}$ given $\bm{\hat{\Gb}}$ and $\bm{\hat{\s}}$ is then a product of likelihoods for $\bm{\pi}$ given each $\hat{\Gb}_i$ and $\hat{\s}_i$. Since the likelihoods are Gaussian and the priors are mixtures of Gaussians, the posterior distribution of $\Gb_i$ given $\hat{\Gb}_i$ and $\bm{\hat{\pi}}$ is analytically tractable, where the mixing proportions $\bm{\hat{\pi}}$ are maximum likelihood estimates. The posterior mean of $\Gb_i$ is taken to be the point estimate of $\Gb_i$, resulting in a shrinkage procedure.\bigskip\\
The flexibility and computational efficiency of ASH is immediately obvious, which motivates us to extend this method to the more complex wavelet settings described here. As an Empirical Bayes procedure, ASH borrows information across all available data and is extremely data adaptive. At the same time, the key advantage of ASH over other Empirical Bayes procedures such as EbayesThresh [Johnstone \& Silverman (2005)] and the commonly used ``spike-and-slab'' prior (is there a name for this?) variable variances are allowed as input, a feature which latter methods lack. This is also the primary reason we can use ASH as the only shrinkage procedure in both the problems described above. Furthermore, Gaussian mixture priors can effectively approximate any unimodal distribution, allowing for a more flexible prior than many other procedures. In tackling these two specific problems, we can speed up the entire signal denoising problem via likelihood approximations (discussed in the next section), allowing our method to be computationally efficient without compromising accuracy. Typical wavelet methods require the specification of a primary resolution level as a ``tuning'' parameter, which could substantially influence the accuracy of the method. However, applying ASH to every resolution in the wavelet transformationwe spares us the need for such a parameter. For an in-depth understanding of the original motivations and applications of ASH, see Stephens (??). With the main shrinkage method accounted for, we will proceed to explore the two aforementioned problems in detail. Note that one only needs to supply a vector of estimates $\hat{\Gb}$ and their standard errors se$(\hat{\Gb})$ to ASH to obtain posterior estimates; hence, we will focus on obtaining these estimates and standard errors when describing our methods.

@@

with $\sum_l \pi_l^{(j)}=1$, for $j=1,...,J$. $\pi_i^{(j)}$ and $(s_i^{(j)})^2$ are hyperparameters that are shared between coefficients in the same resolution level, and $m$ is the number of mixture components. In ASH, $m$ is usually chosen so that it is large enough to reasonably approximate any unimodal distribution, while maintaining an acceptable computational speed. Note that, for notational convenience, we have switched to a double index following standard wavelet convention for indices: here $j=1,...,J$ is the resolution level, and $k=0,...,2^j-1$ is the location within each resolution level $j$. By applying the inverse wavelet transform on the posterior mean $\hat{\bm{\Ga}}$ that ASH produces, one can obtain the posterior mean of $\bm{\mu}$, which serves as an estimate that minimizes the MSE. One could also construct credible bands using the posterior variances.\bigskip\\
While many existing methods do a good job of signal denoising in the presence of i.i.d. (Gaussian) errors, heteroskedastic (but still independent) errors present a different challenge. One key obstacle when dealing with heteroskedastic errors is that the likelihood for $\bm{\Ga}$ does not necessarily factorize. Given the true variances and hence the likelihood, one could compute the posterior distribution using some specified prior. However, this would be computationally cumbersome, and may not be feasible when extended to multiple signals or higher dimensions. A suitable prior might also be difficult to find. As such, one key aspect of our approach is to treat the wavelet coefficients as if they were independent, so that the true likelihood is approximated by a composite likelihood (see Silverman (1999) for an example where this is done).\bigskip\\

@@

where $\Sigma=\textrm{diag}(\s_1,...,\s_T)$ is a diagonal matrix. For the rest of this subsection, we will use the non-decimated wavelet transform (NDWT) instead of the standard DWT to achieve translation invariance. This in turn reduces the presence of artifacts, which often occur near discontinuities in the underlying signal (see eg. Coifman \& Donoho (1995)). Now let $W$ be the matrix associated with the NDWT for a given wavelet basis, so that

@@

where $P(d_{jk}|\Ga_{jk})=\phi(d_{jk};\Ga_{jk},\tilde{\Sigma}_{(jk,jk)})$. Note that there are $n$ coefficients at each resolution level instead of $2^j$ for resolution $j$ due to the NDWT. Here $\phi$ denotes the Gaussian density function. Since $\Sigma$ is diagonal, it is easy to see that $\tilde{\Sigma}_{(jk,jk)}=\sum_{i=1}^T \Sigma_{ii}W_{jk,i}^2$. In our method, we use ASH to assign independent priors to $\Ga_{jk}$ as with \eqref{eq:ashprior}. %ASH also allows us to estimate the $\pi_i$'s via a maximum likelihood approach and fix the $\s_i$'s by using a grid of values. This particular choice of prior specification is extremely flexible, being able to approximate any symmetrical zero-centered distribution, as well as allowing for a relatively easy estimation of the hyperparameters through empirical Bayes. See ASH for more details regarding this choice of prior (?)

@@

Although we have assumed that the true variances were known for mean estimation, the problem of variance estimation itself is a non-trivial one. Here we make the reasonable assumption that the variance function is also spatially structured, in addition to the mean function. The approach we took is similar in spirit to that of Cai \& Wang (2008), since we make use of wavelet decomposition as well. However, while they use first order differences, we look at the squared residuals. As such, we need a reasonably accurate estimator of the mean function to form sensible residuals. For clarity of presentation, we will first assume that the true mean function is known. Define

@@

Now that we can now estimate $\bm{\mu}$ given $\bm{\s}^2$ and vice versa, a natural procedure for jointly estimating $\bm{\mu}$ and $\bm{\s}^2$ is an iterative one. That is, we estimate $\bm{\mu}$ given some initial estimate of $\bm{\s}^2$, then estimate $\bm{\s}^2$ given the previous estimate of $\bm{\mu}$, and iterate the process. In our simulations, we have found that two cycles are usually sufficient to produce relatively accurate estimates. To start off the algorithmm, an initial estimate would be the square of first order differences between adjacent points, as discussed in previous variance estimation papers (eg Cai \& Wang (2008)). This estimator has the property that it is approximately unbiased for the variance at the two corresponding points, provided that the mean and the variance function is smooth enough. To be more specific, the initial variance estimates are defined by

@@

where $Y_0\equiv Y_n$ and $Y_{n+1}\equiv Y_1$, due to periodicity assumptions in wavelet methods. In summary, our approach for joint mean and variance estimation can be described in three steps:

@@

\item Using squared first order differences \eqref{eq:initial var est} as estimates of $\bm{\s}^2$, estimate $\bm{\mu}$ as if $\bm{\s}^2$ was known (see above for the case when $\bm{\s}^2$ is known)
\item Given the estimate of $\bm{\mu}$, take squared residuals and treat those as ``observations'' for $\bm{\s}^2$. Next, project them into wavelet space and apply shrinkage to the wavelet coefficients, before projecting them back into data space to obtain an estimate of $\bm{\s}^2$.

@@

Next we turn to another popular nonparametric regression problem and demonstrate how we could make use of ASH. Specifically, we also assume an underlying signal $\Gl_i$, $i=1,...,n$ with $n=2^J$ for some positive integer $J$. Further assume that each data point $Y_i$, $i=1,...,n$ is realized from a Poisson distribution with mean $\Gl_i$ ie. $Y_i=\textrm{Pois}(\Gl_i)$. Our goal is to recover the true intensity $\bm{\Gl}$ as accurately as possible, given $Y_k$. To make our approach easier to understand and relate it to wavelet-based methods, we first summarize the data in a recursive manner (see Kolaczyk (1999) and Nowak (1998):

@@

for resolution $j=0,...,J-1$ and location $k=0,...,2^j-1$, so that we are looking at the sum of adjacent pairs of ``points'' as we move to coarser resolutions, where a point is already itself a sum of the data points. Another way to define the $Y$'s is

@@

where the factorization is due to the recursive definition above. Note that $Y_{00}|\Gl_{00}\sim \textrm{Pois}(\Gl_{00})$. For any given $j,k$, $Y_{jk}$ is a sum of two independent Poisson random variables, and is itself a Poisson random variable. Hence

@@

Whereas Kolaczyk (1999) sets a conjugate shrinkage prior (a mixture of a point mass at 0.5 and a uniform distribution, similar to the ``spike and slab'' prior for the Gaussian case) for the parameters $\frac{\Gl_{j+1,2k}}{\Gl_{jk}}$ and uses a Binomial likelihood, we felt that the two-component mixture prior is too inflexible. As such, we work with the parameters $\Ga_{sl}$ as defined in \eqref{eq:poisson wc} and approximate the logit-Binomial (is there such a thing?) likelihoods by Gaussian likelihoods. Furthermore, these parameters can be more easily connected to the true wavelet coefficients since they consist of (log) differences rather than ratios.\bigskip\\

@@

where $S$ is the number of successes in the binomial setup, and $F=N-S$ is the number of failures (in our case the even locations would be successes and odd locations failures). The MLE has one notable drawback however, in that it cannot deal well with extreme cases. This would in turn affect the likelihood approximation in a non-negligible way. To improve the accuracy of the approximation at the endpoints of the binomial distribution (ie. when we have $\hat{p}=0$ or $N$, so that $\hat{\Ga}=-\infty$ or $\infty$ respectively), we instead aim to find a suitable estimator $\hat{\Ga}$ for $\Ga$. Next, we approximate the distribution of $\hat{\Ga}$ by a Gaussian distribution, so that the likelihood for $\Ga$ is automatically Gaussian, given by $\phi(\Ga;\hat{\Ga},\mathbb{V}(\hat{\Ga}))$. To that end, our final estimator will still be the MLE, but with Tukey's modification (Gart \& Zweifel (1967)) at the endpoints where $S=\{0,N\}$. Specifically, the estimator for $\Ga$ and its standard error are given by

@@

Finally, the Delta method is used to reconstruct the signal in the original space from the posterior distribution of the $\Ga$'s (which are in the logit space). For full details on the reconstruction process as well as the Delta method, see Appendix \ref{app:reconstruction}.

@@

In all the scenarios presented in Tables \ref{table:hetero_v2_1} - \ref{table:hetero_bump_3}, it is obvious that the mean estimates deviate substantially from the true mean if the heteroskedastic nature of the variance functions is not taken into account. We also note that our method can nearly achieve the ``optimal'' MISE (ie the MISE for the case when our method is given the true variance) for variance function $V_2$, but not for the other. This is likely because function $V_2$ is relatively smooth, and hence is easier to estimate than the extremely jumpy and peaky Bumps function. Furthermore, the MISEs of our method are often smaller than the ones produced by TI-thresholding even when TI-thresholding is supplied the true variance function, where the latter procedure has been shown to be asymptotically optimal (see Gao (1997)). Hence, the finite sample performance of our method is guaranteed to a certain extent through these simulation studies. At the same time, using the variance function estimated by our method instead of running MAD also improves the performance of TI-thresholding, demonstrating the ability of our method to estimate the variance function relatively accurately. However, our method is still superior to TI-thresholding no matter the type of variance estimate that TI-thresholding uses in these simulation studies.\bigskip\\

@@

In the first scenario, the number of data points is not a power of two, nor are the point equally spaced. To deal with these complications, we adapted and modified the standard symmetric extension procedure commonly adopted in wavelet settings. We first mirrored the data about the right edge and extract the first $2^{\lfloor\log_2(2n)\rfloor}$ sample points. This ensures that the number of data points in the new ``dataset'' is a power of two, and the mean curve would be continuous at the right edge. To further ensure that the input to the Gaussian denoising method is periodic, we then reflected the new dataset about the right edge and used this as the final input. To obtain the original mean and variance functions, we extracted the first $n$ points from the outputs (mean and variance) of our denoising technique. Since the data points are not evenly spaced, we took the simplest approach and applied our method treating the observations as if they were evenly spaced. This approach is not only intuitively appealing, but can also be considered a formal treatment of unequally spaced data in traditional wavelet settings (see Sardy et al. (1999)). Evaluation of MSE at the 201 equally spaced points is then based on simple linear interpolation between the estimated points. Tables \ref{table:mfvb_comp} displays the MSEs over 100 independent runs for each scenario.

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & (0.01,3) & (1/8,8) & (1/128,128) \\
  \hline
SMASH & 1.00 & 1.00 & 1.00 \\
  BMSM & 1.55 & 1.58 & 1.21 \\
  BMMIM & 1.32 & 1.86 & 2.06 \\
  Haar-Fisz & 1.20 & 1.39 & 1.31 \\
  Anscombe, universal & 1.55 & 1.88 & 1.39 \\
  Haar thresholds & 1.35 & 1.38 & 1.23 \\
   \hline
\end{tabular}
\caption{Relative MSE of various methods for the Bumps function for Poisson data}
\label{table:pois_bumps}
\end{table}

@@

In this paper we have briefly introduced the adaptive shrinkage method ASH; while it was originally developed in the setting of FDR control for multiple comparisons, we have illustrated its usage as part of two wavelet denoising techniques. Both applications discussed in this paper relax the standard assumption of i.i.d. Gaussian noise, and are thus challenging tasks. Through these applications we are able to demonstrate the flexibility and accuracy of the shrinkage method, demonstrating its potential in many other applications.\bigskip\\
In both the aforementioned applications, our software allows users to easily obtain point estimates for the mean function as well as approximate credible intervals as a measure of uncertainty. In addition, we can also estimate the variance function if desired, which can be used to provide frequentist confidence intervals for other forms of mean estimation. To the best of our knowledge, there is no readily available software which implements joint mean and variance estimation in the wavelet literature. We have also confirmed through simulations that our method is relatively robust to simple forms of autocorrelation between the errors (details needed). In the case of Poisson regression, we improved upon the conjugate Beta priors used in conjunction with the binomial likelihoods by utilizing ASH as the shrinkage procedure, which allows for more flexibility and precision. %In addition, we have proposed a new procedure to assess the performance of a given method on real datasets without knowing the underlying truth, under the assumption that the data follow a Poisson distribution. 
In both the applications, one further advantage of both our methods is that there is no tuning parameter other than the type of wavelet basis used. On the other hand, the primary resolution level in almost all of the other wavelet-based methods actually affects their performance in varying degrees, depending on the underlying mean and/or variance function. Hence, our fully adaptive procedures allow users to easily apply them to any given dataset, depending on context.\bigskip\\
